---
title: "Simple Probability"
output:
  html_document:
    df_print: paged
---

# Introduction

This is to explain some very simple probability & statistics. Most of you already know these things. So this is for absolute beginners.

# Basic probability

Lets assume simple coin. It has a two faces and if you flip the coin it will give  result either **HEAD** or **TAIL**

Given that all outcomes are equally likely, we can compute the probability of a TAIL using the formula:
</br>
$$ \text{Probabillity} = \frac{\text{Number of favourable outcome}}{\text{Number of possible eaqually like outcomes}} $$
</br>
So probability of result HEAD is 0.5 or 1/2 and probability of result TAIL is 0.5 or 1/2
Lets assume we flip the coin 10 times. We can use R to simulate the results.

1. we create the coin & assign 0 & 1 (for HEAD & TAIL)

```{r}
coin <- c(0,1)

```
2. flip the coin 10 times and store the results in **result**

```{r}
result <- sample(coin,10, replace = TRUE)

```
Lets see the result

```{r}
result
```
We can show as a table

```{r}
table(result)
```


Bar chart of this results

```{r}
library(lattice)
barchart(table(result), horizontal = FALSE)

```

## Probability Density Function

The Probability Density Function is

```{r}
densityplot(result)
```
