---
title: "Simple Probability"
output:
  html_document:
    df_print: paged
---


# Introduction

This is to explain some very simple probability & statistics. Most of you already know these things. We can use this to revise some *basic probability* concepts.

# Basic probability

The probababilty of perticuar event can be described as:

$$ \text{Probabillity} = \frac{\text{Number of favourable outcome}}{\text{Number of possible eaqually like outcomes}} $$

Lets assume simple coin. It has a two faces and if you toss the coin it will give result either **HEAD** or **TAIL**

The probability of the coin landing H or HEAD is $P(H)$
And the probability of the coin landing T or TAIL is $P(T)$

Given that all outcomes are equally likely, we can compute the probability of a landing HEAD using the formula:

$$P(H) = \frac{\text{Number of favourable outcome}}{\text{Number of possible eaqually like outcomes}} = \frac{1}{2} $$

Similarly probability of landing TAIL is $\frac{1}{2}$.

Probability is Just a Guide
Probability does not confirm us exactly what will happen, it is just a guide.

Example: toss a coin 10 times, how many Heads will come up?
According to the Probability of the coin landing Head is ½ chance, so we can expect 5 Heads.
But when we actually try it we might get 3 heads, or  4 heads ... or anything really.

Lets simulate this using some R script


1. we create the coin & assign "H" & "T" (for HEAD & TAIL)

```{r coin}
coin <- c("H","T")

```
2. toss the coin 10 times and store the results in **result**

```{r}
result <- sample(coin,10, replace = TRUE)

```
Lets see the result

```{r}
result
```
We can show as a table

```{r}
table(result)
```
Using dplyr & tidyr packages we can arrange the resutls nicely to show table & also include the measured probability of "T" and "H" occurances.

Lets load the required libraries

```{r message= FALSE, warning=FALSE}

library(tidyr)
library(dplyr)
library(ggplot2)
library(gridExtra)

```

Arrange the results


```{r}

df_total <- data.frame(result) %>% count(result)

```

Lets see the `df_total`
```{r}
df_total

```




with above results we can see TAIL `r df_total$T` times and HEAD `r df_total$H` times.

Using `spread` & `gather` in tidyr and `mutate` &  `select` in dplyr package we can add additional columns to show percentage of each "T" & "H"

```{r}

df_prob <- spread(df_total,result,n) %>% 
  mutate(prob_H = H/(H+T) , prob_T = T/(H+T) ) %>% 
  select(prob_H, prob_T) %>% gather(desc,prob,prob_H:prob_T)
df_prob
```




Bar chart of this results

```{r}
ggplot(df_total, aes(result,n, fill = result)) + geom_bar(stat = "identity") + labs(title = "Tossing a coin 10 times") +theme(legend.position = "none")
ggplot(df_prob, aes(desc,prob, fill = desc)) + geom_bar(stat = "identity") + labs(title = "Measured probability of Tossing a coin 10 times", x = "") + theme(legend.position = "none")

```

**What is the conclusion?**

> seems number of HEAD or TAIL frequencies are not equal?

Probability of HEAD $P(H)$

Probability of TAIL $P(T)$

In here why $P(H) = P(T) = \frac{1}{2}$ can not be observed.

> Probability is Just a Guide
> Probability does not confirm us exactly what will happen


## toss Coin function

To demonstrate that we can create **tossCoin** function & **plotCoin** function to plot the results nicely.

```{r}
tossCoin <- function(numToss = 10){
  coin <- c("H","T")
  result <- sample(coin,numToss, replace = TRUE)
  df_total <- data.frame(result) %>% count(result)
  

}

plotCoin <- function(df){
  number <- sum(df$n)
  
  df_prob <- spread(df,result,n) %>% 
  mutate(prob_H = H/(H+T) , prob_T = T/(H+T) ) %>% 
  select(prob_H, prob_T) %>% gather(desc,prob,prob_H:prob_T)
  
  plot1 <- ggplot(df, aes(result,n, fill = result)) + geom_bar(stat = "identity") + labs(title = paste0("Tossing a coin ", number," times")) +theme(legend.position = "none")
  plot2 <- ggplot(df_prob, aes(desc,prob, fill = desc)) + geom_bar(stat = "identity") + labs(title = paste0("Measured probability of Tossing a coin ", number," times"), x = "") + theme(legend.position = "none")
  
  grid.arrange(plot1, plot2)
  
}

```


Lets see the results by tossing coin 100 times

```{r}
df100 <- tossCoin(100)
df100
plotCoin(df100)

```

May be 1000 times


```{r}
df1000 <- tossCoin(1000)
df1000
plotCoin(df1000)


```

