---
title: "Simple Probability"
output:
  html_document:
    df_print: paged
---

# Introduction

This is to explain some very simple probability & statistics. Most of you already know these things. So this is for absolute beginners.

# Basic probability

Lets assume simple coin. It has a two faces and if you flip the coin it will give result either **HEAD** or **TAIL**

Given that all outcomes are equally likely, we can compute the probability of a TAIL using the formula:
</br>
$$ \text{Probabillity} = \frac{\text{Number of favourable outcome}}{\text{Number of possible eaqually like outcomes}} $$
</br>

So probability of result HEAD is 0.5 or 1/2 and probability of result TAIL is 0.5 or 1/2

Probability of an event (either HEAD or TAIL) $P(A) = \frac{1}{2}$ 

Now flip the coin 10 times. We can use R to simulate the results.

1. we create the coin & assign 1 & 0 (for HEAD & TAIL)

```{r}
coin <- c(1,0)

```
2. flip the coin 10 times and store the results in **result**

```{r}
result <- sample(coin,10, replace = TRUE)

```
Lets see the result

```{r}
result
```
We can show as a table

```{r}
table(result)
```


Bar chart of this results

```{r}
library(lattice)
barchart(table(result), horizontal = FALSE, xlab = c("TAIL", "HEAD"))

```

**What is the conclusion?**

> seems number of HEAD or TAIL frequencies are not equal?

Probability of HEAD $P(H)$

Probability of TAIL $P(T)$

In here why $P(H) = P(T) = \frac{1}{2}$ can not be observed.

Because we need to increase the number of coin flips to see that.

## Flip coin function

To demonstrate that we can create **flipCoin** function
```{r}
flipCoin <- function(numFlips = 10){
  coin <- c(1,0)
  sample(coin, numFlips, replace = TRUE)

}


```

Lets see the results by fliping coin 100 times

```{r}
barchart(table(flipCoin(100)), horizontal = FALSE, xlab = c("TAIL", "HEAD"))
```

May be 1000 times


```{r}
barchart(table(flipCoin(1000)), horizontal = FALSE, xlab = c("TAIL", "HEAD"))
```

## Probability Density Function

The Probability Density Function is

```{r}
densityplot(result)
```
