---
title: "Simple Probability"
output:
  html_document:
    df_print: paged
---

# Introduction

This is to explain some very simple probability & statistics. Most of you already know these things. So this is for absolute beginners.

# Basic probability

The probababilty of perticuar event can be described as:

$$ \text{Probabillity} = \frac{\text{Number of favourable outcome}}{\text{Number of possible eaqually like outcomes}} $$

Lets assume simple coin. It has a two faces and if you toss the coin it will give result either **HEAD** or **TAIL**

The probability of the coin landing H or HEAD is $P(H)$
And the probability of the coin landing T or TAIL is $P(T)$

Given that all outcomes are equally likely, we can compute the probability of a landing HEAD using the formula:

$$P(H) = \frac{\text{Number of favourable outcome}}{\text{Number of possible eaqually like outcomes}} = \frac{1}{2} $$

Similarly probability of landing TAIL is $\frac{1}{2}$.

Probability is Just a Guide
Probability does not tell us exactly what will happen, it is just a guide

Example: toss a coin 10 times, how many Heads will come up?
According to the Probability of the coin landing Head is ½ chance, so we can expect 5 Heads.
But when we actually try it we might get 3 heads, or  4 heads ... or anything really.

Lets simulate this using some R script

1. we create the coin & assign 1 & 0 (for HEAD & TAIL)

```{r}
coin <- c(1,0)

```
2. toss the coin 10 times and store the results in **result**

```{r}
result <- sample(coin,10, replace = TRUE)

```
Lets see the result

```{r}
result
```
We can show as a table

```{r}
table(result)
```


Bar chart of this results

```{r}
library(lattice)
barchart(table(result), horizontal = FALSE, xlab = c("TAIL", "HEAD"))

```

**What is the conclusion?**

> seems number of HEAD or TAIL frequencies are not equal?

Probability of HEAD $P(H)$

Probability of TAIL $P(T)$

In here why $P(H) = P(T) = \frac{1}{2}$ can not be observed.

Because we need to increase the number of coin flips to see that.

## Flip coin function

To demonstrate that we can create **flipCoin** function
```{r}
flipCoin <- function(numFlips = 10){
  coin <- c(1,0)
  sample(coin, numFlips, replace = TRUE)

}


```

Lets see the results by fliping coin 100 times

```{r}
barchart(table(flipCoin(100)), horizontal = FALSE, xlab = c("TAIL", "HEAD"))
```

May be 1000 times


```{r}
barchart(table(flipCoin(1000)), horizontal = FALSE, xlab = c("TAIL", "HEAD"))
```

## Probability Density Function

The Probability Density Function is

```{r}
densityplot(result)
```
