---
title: "Simple Probability"
output:
  html_document:
    df_print: paged
---


# Introduction

This is to explain some very simple probability & statistics. Most of you already know these things. We can use this to revise some *basic probability* concepts.

# Basic probability

The probababilty of perticuar event can be described as:

$$ \text{Probabillity} = \frac{\text{Number of favourable outcome}}{\text{Number of possible eaqually like outcomes}} $$

Lets assume simple coin. It has a two faces and if you toss the coin it will give result either **HEAD** or **TAIL**

The probability of the coin landing H or HEAD is $P(H)$
And the probability of the coin landing T or TAIL is $P(T)$

Given that all outcomes are equally likely, we can compute the probability of a landing HEAD using the formula:

$$P(H) = \frac{\text{Number of favourable outcome}}{\text{Number of possible eaqually like outcomes}} = \frac{1}{2} $$

Similarly probability of landing TAIL is $\frac{1}{2}$.

Probability is Just a Guide
Probability does not confirm us exactly what will happen, it is just a guide.

Example: toss a coin 10 times, how many Heads will come up?
According to the Probability of the coin landing Head is ½ chance, so we can expect 5 Heads.
But when we actually try it we might get 3 heads, or  4 heads ... or anything really.

Lets simulate this using some R script


1. we create the coin & assign 1 & 0 (for HEAD & TAIL)

```{r}
coin <- c(1,0)

```
2. toss the coin 10 times and store the results in **result**

```{r}
result <- sample(coin,10, replace = TRUE)

```
Lets see the result

```{r}
result
```
We can show as a table

```{r}
table(result)
```
Using dplyr & tidyr packages we can arrange the resutls nicely also include the percentage of "T" and "H" occurances.

Lets load the required libraries

```{r message= FALSE}

library(tidyr)
library(dplyr)

```

Arrange tthe results


```{r}

df <- data.frame(result)
df <- spread(count(df,result),result,n)
names(df) <- c('T', 'H')

```

Lets see the `df`

```{r}
df
```

with above results we can see TAIL `r df$T` times and HEAD `r df$H` times.

Using mutate in dplyr package we can add additional columns to show percentage of each "T" & "H"

```{r}

df <- mutate(df, percT = T/(T+H), percH = H/(T+H))
df

```




Bar chart of this results

```{r}
barplot(c(df$percT, df$percH), names.arg = c("TAIL", "HEAD"), ylab = "%", col = c("lightblue","lightpink"))

```

**What is the conclusion?**

> seems number of HEAD or TAIL frequencies are not equal?

Probability of HEAD $P(H)$

Probability of TAIL $P(T)$

In here why $P(H) = P(T) = \frac{1}{2}$ can not be observed.

> Probability is Just a Guide
> Probability does not confirm us exactly what will happen


## toss Coin function

To demonstrate that we can create **tossCoin** function & **plotCoin** function to plot the results nicely.

```{r}
tossCoin <- function(numToss = 10){
  coin <- c(1,0)
  result <- sample(coin, numToss, replace = TRUE)
  df <- data.frame(result)
  df <- spread(count(df,result),result,n)
  names(df) <- c('T', 'H')
  df <- mutate(df, percT = T/(T+H), percH = H/(T+H))
  return(df)

}

plotCoin <- function(df){
  barplot(c(df$percT, df$percH), names.arg = c("TAIL", "HEAD"), ylab = "%", col = c("lightblue","lightpink"), ylim = c(0,1))
  
}

```


Lets see the results by tossing coin 100 times

```{r}
df100 <- tossCoin(100)
df100
plotCoin(df100)

```

May be 1000 times


```{r}
df1000 <- tossCoin(1000)
df1000
plotCoin(df1000)


```

