---
title: "Simple Probability"
author: Sudarshana A
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message= FALSE, warning=FALSE)
```


# Introduction

This is to explain some very simple probability &  using R scripts to demonstrate the probability concept. Most of you already know these things. We can use this to revise some *basic probability* concepts.

# Basic probability

The probababilty of perticuar event can be described as:

$$ \text{Probabillity} = \frac{\text{Number of favourable outcome}}{\text{Number of possible eaqually like outcomes}} $$

Lets assume simple coin. It has a two faces and if you toss the coin it will give result either **HEAD** or **TAIL**

The probability of the coin landing H or HEAD is $P(H)$

And the probability of the coin landing T or TAIL is $P(T)$

Given that all outcomes are equally likely, we can compute the probability of a landing HEAD using the formula:

$$P(H) = \frac{\text{Number of favourable outcome}}{\text{Number of possible eaqually like outcomes}} = \frac{1}{2} $$

Similarly probability of landing TAIL is $\frac{1}{2}$.

Probability is *Just a Guide*.

Probability does not confirm us exactly what will happen, it is just a guide.

Example: toss a coin 10 times, how many Heads will come up?
According to the Probability of the coin landing Head is ½ chance, so we can expect 5 Heads.
But when we actually try it we might get 3 heads, or  4 heads ... or anything really.

Lets simulate this using some R script


1. we create the coin & assign "H" & "T" (for HEAD & TAIL)

```{r coin}
coin <- c("H","T")

```
2. toss the coin 10 times and store the results in **result**

```{r result}
result <- sample(coin,10, replace = TRUE)

```
Lets see the result

```{r}
result
```
We can show as a table

```{r}
table(result)

```
Lets convert the table to data frame.


```{r}

df_total <- data.frame(table(result))

```

Lets see the `df_total`
```{r}
df_total

```


with above results we can see HEAD `r df_total$Freq[df_total$result == "H"]` times and TAIL `df_total$Freq[df_total$result == "T"]` times.

Using `prop.table()` we can add additional columns to show probability of each "H" & "T"

```{r}
df_prob <- cbind(df_total, prop.table(df_total$Freq))
names(df_prob) <- c("result","Freq", "probability")
df_prob
```


To show the results graphically we can generate bar plot using `barplot` in base R



```{r}
barplot(df_prob$Freq, names.arg = df_total$result, main = "Head & Tail count of Tossing a coin 10 times")
barplot(df_prob$probability, names.arg = df_total$result, main = "Probability of Tossing a coin 10 times")

```

**What is the conclusion?**

> seems number of HEAD or TAIL frequencies are equal/ not equal?

Probability of HEAD $P(H)$

Probability of TAIL $P(T)$

In here $P(H) = P(T) = \frac{1}{2}$ may be/may not be observed.

> Probability is Just a Guide
> Probability does not confirm us exactly what will happen


## toss Coin function

To demonstrate that we can create **tossCoin** function & **plotCoin** function to plot the results nicely.

We can use `ggplot2` package to create plots and `gridExtra` package to arrange the plots.

```{r}
library(ggplot2)
library(gridExtra)
```

Lets write the two functions.

```{r funcCoin}
tossCoin <- function(numToss = 10){
  coin <- c("H","T")
  result <- sample(coin,numToss, replace = TRUE)
  df_total <- data.frame(table(result))
  

}

plotCoin <- function(df){
  number <- sum(df$Freq)
  
  
  plot1 <- ggplot(df, aes(result,Freq, fill = result)) + geom_bar(stat = "identity") + labs(title = paste0("Tossing a coin ", number," times")) +theme(legend.position = "none")
  plot2 <- ggplot(df, aes(result, prop.table(df$Freq), fill = result)) + geom_bar(stat = "identity") + labs(title = paste0("Measured probability of Tossing a coin ", number," times"), y = "Probability") + theme(legend.position = "none")
  
  grid.arrange(plot1, plot2)
  
}

```


Lets see the results by tossing coin 50 times

```{r 50Coin}
df50 <- tossCoin(50)
df50
plotCoin(df50)

```

May be 1000 times


```{r 1000Coin}
df1000 <- tossCoin(1000)
df1000
plotCoin(df1000)


```


> So if we toss the coin large number of times, the probability of each occurance is approximately/equal to the theoritical probability.

Lets try some big number...

```{r millionCoin}
df1million <- tossCoin(1000000)
df1million
plotCoin(df1million)

```
***

## Can you think of similar procedure for toss a Dice.

In here instead of two outcomes we have 6 outcomes.



## toss Dice function

To demonstrate that we can create **tossDice** function & **plotDice** function similar to toss Coin.

```{r funcDice}
tossDice <- function(numToss = 10){
  dice <- c(1:6)
  result <- sample(dice, numToss, replace = TRUE)
  df_total <- data.frame(table(result))
  
}

plotDice <- function(df){
  number <- sum(df$Freq)
  
 
  plot1 <- ggplot(df, aes(result,Freq, fill = result)) + geom_bar(stat = "identity") + labs(title = paste0("Tossing a dice ", number," times")) +theme(legend.position = "none")


  plot2 <- ggplot(df, aes(result, prop.table(df$Freq), fill = result)) + geom_bar(stat = "identity") + labs(title = paste0("Probability of Tossing a dice ", number," times"), y = "Probability") + theme(legend.position = "none")
  
  grid.arrange(plot1, plot2)

}

```


Lets see the results by tossing coin 50 times

```{r 50Dice}
df50 <- tossDice(50)
df50
plotDice(df50)

```

May be 1000 times


```{r 1000Dice}
df1000 <- tossDice(1000)
df1000
plotDice(df1000)


```

Dice can be tossed 1 million times


```{r millionDice}
df1million <- tossDice(1000000)
df1million
plotDice(df1million)

```

as a proportional table

```{r}
cbind(df1million, prop.table(df1million$Freq))


```


> We can see if the number of times tossing dice increases the probablity of each occurance is approximating to $\frac{1}{6} = 0.1666667$


